---
title: "Developing Data Projects Course Project 1 - R Markdown and Leaflet"
author: "Daniel DeWaters"
date: "12/11/2019"
output: html_document
---
# Synopsis

Baltimore is enclosed in a big circle, so I thought is would be creative to make a leaflet map that took in some GPS coordinates that makes a smiley face when displayed. I got the coordinates by dropping pins in google maps around baltimore and copying/pasting the points into a data frame. I then built the leaflet map by plotting the eye points using addMarkers() and making a "smile" line across coordinate points using addPolyLines()

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
# Load libraries
library(leaflet)
library(sp)
```

# Code

```{r points}
# Make coordinates for eye points
lat <- c( 39.353197,  39.360452)
lng<-  c(-76.674711, -76.563840)

eye_markers = data.frame(lat, lng)

# Make coordinates for smile line
lat <- c( 39.306558,  39.283548,  39.280657,  39.284773,  39.307927)
lng <- c(-76.697274, -76.666095, -76.612751, -76.575872, -76.534644)

smile_markers = data.frame(lat, lng)

# Convert data frames into spatial points data frames
eye_markers <- SpatialPointsDataFrame(data=eye_markers, coords=eye_markers)
smile_markers <- SpatialPointsDataFrame(data=smile_markers, coords=smile_markers)

# Build leaflet map
map <- 
  leaflet() %>%
  addTiles() %>%
  addMarkers(data=eye_markers, lng=~lng, lat=~lat) %>%
  addPolylines(data=smile_markers, lng=~lng, lat=~lat)

# Display map
map
```